---
title: ""
output: html_document
---

# Example: Cut Diet

Source (Danish)

* Fødevaredata (http://frida.fooddata.dk), version 4, 2019
* Fødevareinstituttet, Danmarks Tekniske Universitet

Source (English)

* Food data (http://frida.fooddata.dk), version 4, 2019
* The Food Institute, Technical University of Denmark


```{r include=FALSE}

options(scipen=999)

library(openxlsx)
library(tidyr)
library(dplyr)
library(lpSolveAPI)
library(janitor)
library(zoo)
library(DT)
library(shiny)
library(lubridate)
```

## Load data

```{r}

temp_dir = tempdir()
temp_dir

```

```{r}
path = "/home/oem/Dropbox/R/macros-on-point/data"
file_path = paste0(path , "/" , list.files(path = path) )
file_path
```

```{r}

unzip(file_path,exdir = tempdir())
list.files(tempdir())
```


## Display example data table

```{r}

english_file = "Frida20190802env3.xlsx"
sheet_name = "Frida_20190802"

df = openxlsx::read.xlsx(xlsxFile = paste0(tempdir(),"/",english_file),sheet=sheet_name)
df %>% 
  head() %>% 
  DT::datatable(options = list(scrollX = TRUE))
```

## Clean data table

```{r}

df_clean = df %>% 
  slice(2:nrow(.)) %>% 
  dplyr::select(-X1) %>% 
  dplyr::rename(Group=X2,
                Name=X3)
df_clean %>% 
  head() %>% 
  DT::datatable(options = list(scrollX = TRUE))
```


## Tidy Data

```{r}

df_basic_tidy = df_clean %>% 
  dplyr::select(Group,
                Name,
                `Energy,.kcal`,
                `Protein,.videnskabelig`,
                `Carbohydrate,.declaration`,
                `Dietary.fiber`,
                `Fat,.total`,
                Alcohol) %>% 
  setNames(.,c("Group","Name","Kcal","Protein","Carbohydrates","Fiber","Fat","Alcohol")) %>% 
  group_by(Group,Name) %>% 
  filter(n()==1) %>% 
  ungroup() %>% 
  gather(Metric,Value,Kcal:Alcohol) %>% 
  dplyr::mutate(Value=as.numeric(Value)) 

df_basic_tidy %>% 
  head() %>% 
  DT::datatable(options = list(scrollX = TRUE))

```


## Spread data

```{r}

df_basic = df_basic_tidy %>% 
  spread(Metric,Value)

df_basic %>% 
  head() %>% 
  DT::datatable(options = list(scrollX = TRUE))

```


## Settings

```{r echo=FALSE}

weight_kg = 87
age = as.integer(floor(difftime(Sys.Date(),dmy("25-11-1986"),units = "weeks")/52))
height = 185

# bmr = 88.362 + (13.397*weight) + (4.799*height) - (5.677*age)
bmr = 10*weight_kg+6.25*height-5*age+5 

protein_per_kg = 2
fibre_per_kg = 1/3
fat_per_kg = 0.6

decided_deficit_kcal = 500

```


## Functions

```{r echo=FALSE}

## Opskriftsfunktion
calculate_contents = function(Dataframe,Names,Grams) {
  return(
Dataframe %>% 
  dplyr::select(Name,Kcal,Protein,Fat,Carbohydrates,Fiber) %>% 
  inner_join(data.frame(Navn=Names,Gram=Gram)) %>% 
  dplyr::mutate(Kcal=Kcal*Gram/100,
                Protein=Protein*Gram/100,
                Fat=Fat*Gram/100,
                Carbohydrates=Carbohydrates*Gram/100,
                Fiber=Fiber*Gram/100))
}
```

## Optimization algorithm

```{r echo=FALSE,results='hide'}

### OBJECT
lprec <- make.lp(0, nrow(df_basic))
lprec.ctrl = lp.control(lprec,sense='min')
set.objfn(lprec, df_basic$Kcal)

### MACROS
add.constraint(lprec, df_basic$Fat, "=", weight_kg*fat_per_kg*7) 
add.constraint(lprec, df_basic$Fiber, ">=", weight_kg*fibre_per_kg*7) 
add.constraint(lprec, df_basic$Protein, "=", weight_kg*protein_per_kg*7) 

### DF SETTINGS
RowNames <- c("Fat", "Fiber", "Protein")
ColNames <- df_basic$Name

solve(lprec)
```

## Results

```{r echo=FALSE}

df.solved <- df_basic %>% 
    dplyr::mutate(Amount=round(100*get.variables(lprec))) %>% 
    filter(Amount>0)

df.solved %>% 
  DT::datatable(options = list(scrollX = TRUE)) %>% 
  formatRound(c('Kcal','Protein','Fat','Carbohydrates','Fiber','Alcohol','Amount'),0)


```


## Total

```{r echo=FALSE}
df_solved.total = 
  df.solved %>% 
  dplyr::mutate(Kcal=Kcal*Amount/100) %>% 
  dplyr::mutate(Protein=Protein*Amount/100) %>% 
  dplyr::mutate(Fat=Fat*Amount/100) %>% 
  dplyr::mutate(Carbohydrates=Carbohydrates*Amount/100) %>% 
  dplyr::mutate(Fiber=Fiber*Amount/100) %>% 
  dplyr::mutate(Alcohol=Alcohol*Amount/100) %>%
  dplyr::mutate(Amount_Per_Day=Amount/7) %>%
  adorn_totals("row")

df_solved.total %>% 
  DT::datatable(options = list(scrollX = TRUE)) %>% 
  formatRound(c('Kcal','Protein','Fat','Carbohydrates','Fiber','Alcohol','Amount','Amount_Per_Day'),0)

```


## Per Day

```{r echo=FALSE}
df_solved.total %>% 
  add_row(Group="Per Day",
          Name="-",
          Kcal=.$Kcal[nrow(.)]/7,
          Protein=.$Protein[nrow(.)]/7,
          Fat=.$Fat[nrow(.)]/7,
          Carbohydrates=.$Carbohydrates[nrow(.)]/7,
          Fiber=.$Fiber[nrow(.)]/7,
          Alcohol=.$Alcohol[nrow(.)]/7) %>% 
  slice(nrow(.)) %>% 
  DT::datatable(options = list(scrollX = TRUE)) %>% 
  formatRound(c('Kcal','Protein','Fat','Carbohydrates','Fiber','Alcohol','Amount','Amount_Per_Day'),0)
```


## Per Kg Per Day

```{r echo=FALSE}

df_solved.total %>% 
  add_row(Group="Per Day",
          Name="-",
          Kcal=.$Kcal[nrow(.)]/7,
          Protein=.$Protein[nrow(.)]/7,
          Fat=.$Fat[nrow(.)]/7,
          Carbohydrates=.$Carbohydrates[nrow(.)]/7,
          Fiber=.$Fiber[nrow(.)]/7,
          Alcohol=.$Alcohol[nrow(.)]/7) %>% 
  add_row(Group="Per Kg/Day",
          Name="-",
          Kcal=.$Kcal[nrow(.)]/weight_kg,
          Protein=.$Protein[nrow(.)]/weight_kg,
          Fat=.$Fat[nrow(.)]/weight_kg,
          Carbohydrates=.$Carbohydrates[nrow(.)]/weight_kg,
          Fiber=.$Fiber[nrow(.)]/weight_kg,
          Alcohol=.$Alcohol[nrow(.)]/weight_kg) %>% 
  slice(nrow(.)) %>% 
  DT::datatable(options = list(scrollX = TRUE)) %>% 
  formatRound(c('Kcal','Protein','Fat','Carbohydrates','Fiber','Alcohol','Amount','Amount_Per_Day'),2)
  
```

## Necessary steps per day

```{r echo=FALSE}

target_kcal = bmr-decided_deficit_kcal

optimized_kcal_per_day = df_solved.total$Kcal[nrow(df_solved.total)]/7

gap = (optimized_kcal_per_day-target_kcal)/7-decided_deficit_kcal

necessary_steps = gap/55*1000
if (necessary_steps<0) {
  msg = "You don't need to walk extra to reach this deficit." 
} else {
  msg = paste0("You will need to walk ",necessary_steps," steps every day.")
}

```

You have requested a calorie deficit of `r decided_deficit_kcal`.

Your BMR is `r bmr`.

Your calorie intake will be `r round(optimized_kcal_per_day)`.

`r msg`.






